\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-05-20 doc/USER_MANUAL/02_getting_started.tex
updates manual using verbatim for better display in markdown

To download the SPECFEM3D\_Cartesian software package, type this:
\begin{lyxcode}
git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
\end{lyxcode}
{\small \par}

We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
limit to the size of the Unix stack.\\

Then, to configure the software for your system, run the \texttt{configure}
shell script. This script will attempt to guess the appropriate configuration
values for your system. However, at a minimum, it is recommended that
you explicitly specify the appropriate command names for your Fortran
compiler and MPI package (another option is to define FC, CC and MPIF90
in your .bash\_profile or your .cshrc file):
\begin{lyxcode}
./configure~FC=ifort~MPIFC=mpif90
\end{lyxcode}
Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.

@@

The SCOTCH library \citep{PeRo96} provides efficent static mapping,
graph and mesh partitioning routines. SCOTCH is a free software package
developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
In case no SCOTCH libraries can be found on the system, the configuration
will bundle the version provided with the source code for compilation.
The path to an existing SCOTCH installation can to be set explicitly
with the option \texttt{-{}-with-scotch-dir}. Just as an example:
\begin{lyxcode}
./configure~FC=ifort~MPIFC=mpif90~-{}-with-scotch-dir=/opt/scotch
\end{lyxcode}
If you use the Intel ifort compiler to compile the code, we recommend
that you use the Intel icc C compiler to compile Scotch, i.e., use:
\begin{lyxcode}
./configure~CC=icc~FC=ifort~MPIFC=mpif90
\end{lyxcode}

>>>>

To download the SPECFEM3D\_Cartesian software package, type this:
\begin{verbatim}
git clone --recursive --branch devel https://github.com/geodynamics/specfem3d.git
\end{verbatim}
%
We recommend that you add {\texttt{ulimit -S -s unlimited}} to your
{\texttt{.bash\_profile}} file and/or {\texttt{limit stacksize
unlimited }} to your {\texttt{.cshrc}} file to suppress any potential
limit to the size of the Unix stack.\\

Then, to configure the software for your system, run the \texttt{configure}
shell script. This script will attempt to guess the appropriate configuration
values for your system. However, at a minimum, it is recommended that
you explicitly specify the appropriate command names for your Fortran
compiler and MPI package (another option is to define FC, CC and MPIF90
in your .bash\_profile or your .cshrc file):
\begin{verbatim}
./configure FC=ifort MPIFC=mpif90
\end{verbatim}
Note that MPI must be installed with MPI-IO enabled because parts of SPECFEM3D perform I/Os through MPI-IO.

@@

The SCOTCH library \citep{PeRo96} provides efficent static mapping,
graph and mesh partitioning routines. SCOTCH is a free software package
developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
In case no SCOTCH libraries can be found on the system, the configuration
will bundle the version provided with the source code for compilation.
The path to an existing SCOTCH installation can to be set explicitly
with the option \texttt{-{}-with-scotch-dir}. Just as an example:
\begin{verbatim}
./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch
\end{verbatim}
If you use the Intel ifort compiler to compile the code, we recommend
that you use the Intel icc C compiler to compile Scotch, i.e., use:
\begin{verbatim}
./configure CC=icc FC=ifort MPIFC=mpif90
\end{verbatim}

### 2015-05-20 doc/USER_MANUAL/03_mesh_generation.tex
updates manual using verbatim for better display in markdown

\item [{materials\_file}] Contains the material associations for each element.
The format is:

\begin{lyxcode}
element\_ID~material\_ID
\end{lyxcode}

@@

\begin{lyxcode}
domain\_ID~material\_ID~rho~vp~vs~Qkappa~Qmu~anisotropy\_flag
\end{lyxcode}

@@

In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
to the different mesh layers. To do so, you will then need to edit the file
called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:

    0 1 vol1 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy
    0 2 vol2 --> syntax: \#material\_domain\_id \#material\_id \#rho \#vp \#vs \#Q\_kappa \#Q\_mu \#anisotropy

(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
to for instance

    2 1 1500 2300 1800 9999.0 9999.0 0
    2 2 1600 2500 20000 9999.0 9999.0 0

@@

\noindent
Then, in the main directory, type
\begin{lyxcode}
{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
\end{lyxcode}
and use
\begin{lyxcode}
{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
\end{lyxcode}
to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
that can be used to investigate mesh quality, e.g. skewness of elements,
and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
The histogram is also printed to the screen. Analyze that skewness
histogram of mesh elements to make sure no element has a skewness
above approximately 0.75, otherwise the mesh is of poor quality (and
if even a single element has a skewness value above 0.80, then you
must definitely improve the mesh). If you want to start designing
your own meshes, this tool is useful for viewing your creations. You
are striving for meshes with elements with `cube-like' dimensions,
e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}

@@

When you are ready to compile, in the main directory type
\begin{lyxcode}
{\small make~xdecompose\_mesh}{\small \par}
\end{lyxcode}
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.

The partitioning is done in serial for now (in the next release we
will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
\begin{lyxcode}
./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
\end{lyxcode}
where
\begin{itemize}
\item \texttt{nparts} is the number of partitions, i.e., the number of cores
for the parallel simulations,
\item \texttt{input\_directory} is the directory which holds all the files
generated by the Python script \texttt{cubit2specfem3d.py} explained
in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
and
\item \texttt{output\_directory} is the directory for the output of this
partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
for each partition. These files will be needed for creating the distributed
databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\end{itemize}
Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
must be placed in the \texttt{input\_directory} folder before running
the program.

@@

\noindent
In the main directory, type
\begin{lyxcode}
{\small make~xmeshfem3D}{\small \par}
\end{lyxcode}
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the main
directory, as most of the binaries of the package.

@@

From http://en.wikipedia.org/wiki/Universal\_Transverse\_Mercator\_coordinate\_system :

@@

@@

\begin{lyxcode}
material\_ID~rho~vp~vs~Q~anisotropy\_flag~domain\_ID
\end{lyxcode}

@@

\begin{lyxcode}
NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
\end{lyxcode}
\end{description}
The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
defines the file which contains the settings of the topography grid
and of the interfaces grids. Topography is defined as a set of elevation
values on a regular 2D grid. It is also possible to define interfaces
between the layers of the model in the same way. The file needs to
define several parameters:
\begin{itemize}
\item The number of interfaces, including the topography. This needs to
be set at the first line. Then, from the bottom to the top of the
model, you need to define the grids with:
\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
\item number of points along $x$ and $y$ direction (NXI and NETA),
\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
and
\item the name of the file which contains the elevation values (in $y$.$x$
increasing order).
\end{itemize}
At the end of this file, you simply need to set the number of spectral
elements in the vertical direction for each layer. We provide a few
models in the {\texttt{EXAMPLES/}} directory. \\

@@

Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
in \texttt{output\_mesher.txt}; this file provides lots of details
about the mesh that was generated. Please note that the mesher suggests
a time step \texttt{DT} to run the solver with. The mesher output
file also contains a table about the quality of the mesh to indicate
possible problems with the distortions of elements. Alternatively,
output can be directed to the screen instead by uncommenting a line
in \texttt{constants.h}:
\begin{lyxcode}
!~uncomment~this~to~write~messages~to~the~screen~

!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT
\end{lyxcode}
To control the quality of the mesh, check the standard output (either
on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
and analyze the skewness histogram of mesh elements to make sure no
element has a skewness above approximately 0.75, otherwise the mesh
is of poor quality (and if even a single element has a skewness value
above 0.80, then you must definitely improve the mesh). To draw the
skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.

>>>>

\item [{materials\_file}] Contains the material associations for each element.
The format is:
\begin{verbatim}
element_ID material_ID
\end{verbatim}

@@

\begin{verbatim}
domain_ID material_ID rho vp vs Qkappa Qmu anisotropy_flag
\end{verbatim}

@@

In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
to the different mesh layers. To do so, you will then need to edit the file
called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
{\small
\begin{verbatim}
    0 1 vol1 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
    0 2 vol2 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
\end{verbatim}
}
(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
to for instance
{\small
\begin{verbatim}
    2 1 1500 2300 1800 9999.0 9999.0 0
    2 2 1600 2500 20000 9999.0 9999.0 0
\end{verbatim}
}

@@

\noindent
Then, in the main directory, type
{\small
\begin{verbatim}
make xcheck_mesh_quality_CUBIT_Abaqus
\end{verbatim}
}
and use
{\small
\begin{verbatim}
./bin/xcheck_mesh_quality_CUBIT_Abaqus
\end{verbatim}
}
to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
that can be used to investigate mesh quality, e.g. skewness of elements,
and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
The histogram is also printed to the screen. Analyze that skewness
histogram of mesh elements to make sure no element has a skewness
above approximately 0.75, otherwise the mesh is of poor quality (and
if even a single element has a skewness value above 0.80, then you
must definitely improve the mesh). If you want to start designing
your own meshes, this tool is useful for viewing your creations. You
are striving for meshes with elements with `cube-like' dimensions,
e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}

@@

When you are ready to compile, in the main directory type
{\small
\begin{verbatim}
make xdecompose_mesh
\end{verbatim}
}
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.

The partitioning is done in serial for now (in the next release we
will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
\begin{verbatim}
./bin/xdecompose_mesh nparts input_directory output_directory
\end{verbatim}
where
\begin{itemize}
\item \texttt{nparts} is the number of partitions, i.e., the number of cores
for the parallel simulations,
\item \texttt{input\_directory} is the directory which holds all the files
generated by the Python script \texttt{cubit2specfem3d.py} explained
in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
and
\item \texttt{output\_directory} is the directory for the output of this
partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
for each partition. These files will be needed for creating the distributed
databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\end{itemize}
Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
must be placed in the \texttt{input\_directory} folder before running
the program.

@@

\noindent
In the main directory, type
{\small
\begin{verbatim}
make xmeshfem3D
\end{verbatim}
}
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the main
directory, as most of the binaries of the package.

@@

From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\\

@@

%

@@

\begin{verbatim}
material_ID rho vp vs Q anisotropy_flag domain_ID
\end{verbatim}

@@

\begin{verbatim}
NEX_XI_BEGIN NEX_XI_END NEX_ETA_BEGIN NEX_ETA_END NZ_BEGIN NZ_END material_ID
\end{verbatim}
\end{description}
The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
defines the file which contains the settings of the topography grid
and of the interfaces grids. Topography is defined as a set of elevation
values on a regular 2D grid. It is also possible to define interfaces
between the layers of the model in the same way. The file needs to
define several parameters:
\begin{itemize}
\item The number of interfaces, including the topography. This needs to
be set at the first line. Then, from the bottom to the top of the
model, you need to define the grids with:
\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
\item number of points along $x$ and $y$ direction (NXI and NETA),
\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
and
\item the name of the file which contains the elevation values (in $y$.$x$
increasing order).
\end{itemize}
At the end of this file, you simply need to set the number of spectral
elements in the vertical direction for each layer. We provide a few
models in the {\texttt{EXAMPLES/}} directory. \\

@@

Mesher output is provided in the \texttt{OUTPUT\_FILES} directory
in \texttt{output\_mesher.txt}; this file provides lots of details
about the mesh that was generated. Please note that the mesher suggests
a time step \texttt{DT} to run the solver with. The mesher output
file also contains a table about the quality of the mesh to indicate
possible problems with the distortions of elements. Alternatively,
output can be directed to the screen instead by uncommenting a line
in \texttt{constants.h}:
\begin{verbatim}
! uncomment this to write messages to the screen
! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
\end{verbatim}
To control the quality of the mesh, check the standard output (either
on the screen or in the \texttt{OUTPUT\_FILES} directory in \texttt{output\_mesher.txt})
and analyze the skewness histogram of mesh elements to make sure no
element has a skewness above approximately 0.75, otherwise the mesh
is of poor quality (and if even a single element has a skewness value
above 0.80, then you must definitely improve the mesh). To draw the
skewness histogram on the screen, type \texttt{gnuplot plot\_mesh\_quality\_histogram.gnu}.

### 2015-05-20 doc/USER_MANUAL/04_creating_databases.tex
updates manual using verbatim for better display in markdown

\noindent
In the main directory, type
\begin{lyxcode}
{\small make~xgenerate\_databases}{\small \par}
\end{lyxcode}
Input for the program is provided through the main parameter file
\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
Please note that \texttt{xgenerate\_databases} must be called directly
from the main directory, as most of the binaries of the package.

@@

2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
each running on 100 cores (see e.g. http://www.schedmd.com/slurmdocs/job\_array.html )

@@

The way we prescribe material property for porous material in SPECFEM3D is as follow:
We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:\\
 rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr\\
where\\
rho\_s = solid density\\
rho\_f = fluid density\\
phi = porosity\\
tort = tortuosity\\
kxx = xx component of permeability tensor\\
kxy = xy,yx components of permeability tensor\\
kyy = yy component of permeability tensor\\
kxz = xz,zx components of permeability tensor\\
kzz = zz component of permeability tensor\\
kappa\_s = solid bulk modulus\\
kappa\_f= fluid bulk modulus\\
kappa\_fr= frame bulk modulus\\
eta\_f = fluid viscosity\\
mu\_fr = frame shear modulus.\\

>>>>

\noindent
In the main directory, type
{\small
\begin{verbatim}
make xgenerate_databases
\end{verbatim}
}
Input for the program is provided through the main parameter file
\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
Please note that \texttt{xgenerate\_databases} must be called directly
from the main directory, as most of the binaries of the package.

@@

2/ submit a single job on 1000 cores to the batch, and in that script create a sub-array of jobs to start 10 jobs,
each running on 100 cores (see e.g. \url{http://www.schedmd.com/slurmdocs/job_array.html})

@@

The way we prescribe material property for porous material in SPECFEM3D is as follow:
We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:
\begin{verbatim}
 rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim} 
where\\
rho\_s = solid density\\
rho\_f = fluid density\\
phi = porosity\\
tort = tortuosity\\
kxx = xx component of permeability tensor\\
kxy = xy,yx components of permeability tensor\\
kyy = yy component of permeability tensor\\
kxz = xz,zx components of permeability tensor\\
kzz = zz component of permeability tensor\\
kappa\_s = solid bulk modulus\\
kappa\_f= fluid bulk modulus\\
kappa\_fr= frame bulk modulus\\
eta\_f = fluid viscosity\\
mu\_fr = frame shear modulus.\\

### 2015-05-20 doc/USER_MANUAL/05_running_the_solver.tex
updates manual using verbatim for better display in markdown

Now that you have successfully generated the databases, you are ready
to compile the solver. In the main directory, type
\begin{lyxcode}
{\small make~xspecfem3D}{\small \par}
\end{lyxcode}
Please note that \texttt{xspecfem3D} must be called directly from
the main directory, as most of the binaries of the package.\\

@@

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
%\begin{lyxcode}
\begin{figure}[htp]
\noindent
\begin{centering}
{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.pdf}
}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%\end{lyxcode}
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type

\begin{lyxcode}
{\small make~xconvolve\_source\_timefunction}{\small \par}
\end{lyxcode}

to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
or the centroid time of the earthquake. The start time of the simulation
is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
rate function is very close to zero when starting the simulation).
To convert to absolute time $t_{\mathrm{abs}}$, set

\begin{lyxcode}
$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
\end{lyxcode}

@@

Each line represents one station in the following format:
\begin{lyxcode}
{\small Station~Network~Latitude~(degrees)~Longitude~(degrees)~Elevation~(m)~burial~(m)~}{\small \par}
\end{lyxcode}
The solver \texttt{xspecfem3D} filters the list of stations in file
\texttt{DATA/STATIONS} to exclude stations that are not located within
the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.

Solver output is provided in the \texttt{OUTPUT\_FILES} directory
in the \texttt{output\_solver.txt} file. Output can be directed to
the screen instead by uncommenting a line in \texttt{constants.h}:
\begin{lyxcode}
!~uncomment~this~to~write~messages~to~the~screen~

!~integer,~parameter~::~IMAIN~=~ISTANDARD\_OUTPUT~
\end{lyxcode}
On PC clusters the seismogram files are generally written to the local
disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
need to be gathered at the end of the simulation.

While the solver is running, its progress may be tracked by monitoring
the `\texttt{\small timestamp{*}}{\small ' files in the ~}\\
{\small{} }\texttt{\small OUTPUT\_FILES}{\small{} directory. These tiny
files look something like this:}{\small \par}
\begin{lyxcode}
Time~step~\#~~~~~~~~~~10000~

Time:~~~~~108.4890~~~~~~seconds~

Elapsed~time~in~seconds~=~~~~1153.28696703911~

Elapsed~time~in~hh:mm:ss~=~~~~~0~h~19~m~13~s~

Mean~elapsed~time~per~time~step~in~seconds~=~~~~~0.115328696703911~

Max~norm~displacement~vector~U~in~all~slices~(m)~=~~~~1.0789589E-02~
\end{lyxcode}
The \texttt{\small timestamp{*}}{\small{} files provide the }\texttt{\small Mean
elapsed time per time step in seconds}{\small , which may be used
to assess performance on various machines (assuming you are the only
user on a node), as well as the }\texttt{\small Max norm displacement
vector U in all slices~(m)}{\small . If something is wrong with the
model, the mesh, or the source, you will see the code become unstable
through exponentially growing values of the displacement and fluid
potential with time, and ultimately the run will be terminated by
the program. You can control the rate at which the timestamp files
are written based upon the parameter }\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}{\small{}
in the }\texttt{\small Par\_file}{\small .}{\small \par}

>>>>

Now that you have successfully generated the databases, you are ready
to compile the solver. In the main directory, type
{\small
\begin{verbatim}
make xspecfem3D
\end{verbatim}
}
Please note that \texttt{xspecfem3D} must be called directly from
the main directory, as most of the binaries of the package.\\

@@

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
\begin{figure}[htp]
\noindent
\begin{centering}
{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.pdf}
}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type
{\small
\begin{verbatim}
make xconvolve_source_timefunction
\end{verbatim}
}
to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
or the centroid time of the earthquake. The start time of the simulation
is $t=-1.5*\texttt{half duration}$ (the 1.5 is to make sure the moment
rate function is very close to zero when starting the simulation).
To convert to absolute time $t_{\mathrm{abs}}$, set
%
\begin{lyxcode}
$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
\end{lyxcode}

@@

Each line represents one station in the following format:
{\small
\begin{verbatim}
Station Network Latitude(degrees) Longitude(degrees) Elevation(m) burial(m)
\end{verbatim}
}
The solver \texttt{xspecfem3D} filters the list of stations in file
\texttt{DATA/STATIONS} to exclude stations that are not located within
the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.

Solver output is provided in the \texttt{OUTPUT\_FILES} directory
in the \texttt{output\_solver.txt} file. Output can be directed to
the screen instead by uncommenting a line in \texttt{constants.h}:
\begin{verbatim}
! uncomment this to write messages to the screen
! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
\end{verbatim}
On PC clusters the seismogram files are generally written to the local
disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
need to be gathered at the end of the simulation.

While the solver is running, its progress may be tracked by monitoring
the `\texttt{\small timestamp{*}}' files in the \texttt{\small OUTPUT\_FILES/} directory. These tiny
files look something like this:
{\small
\begin{verbatim}
Time step #          10000
Time:     108.4890      seconds
Elapsed time in seconds =    1153.28696703911
Elapsed time in hh:mm:ss =     0 h 19 m 13 s
Mean elapsed time per time step in seconds =     0.115328696703911
Max norm displacement vector U in all slices (m) =     1.0789589E-02
\end{verbatim}
}
The \texttt{\small timestamp{*}} files provide the 
\texttt{\small Mean elapsed time per time step in seconds}, which may be used
to assess performance on various machines (assuming you are the only
user on a node), as well as the 
\texttt{\small Max norm displacement vector U in all slices~(m)}. 
If something is wrong with the
model, the mesh, or the source, you will see the code become unstable
through exponentially growing values of the displacement and fluid
potential with time, and ultimately the run will be terminated by
the program. You can control the rate at which the timestamp files
are written based upon the parameter 
\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}
in the \texttt{\small Par\_file}.
